---
title: "Lab3"
author: "Derek Walker"
format:
  html:
    embed-resources: true
---
## 1. Read in the Data

```{r message=F}
if(!require("R.utils")){ install.packages("R.utils")
library("R.utils")}
download.file(
  "https://raw.githubusercontent.com/USCbiostats/data-science-data/master/02_met/met_all.gz",
  destfile = file.path("~", "Downloads", "met_all.gz"),
  method   = "libcurl",
  timeout  = 60
)

met <- data.table::fread(file.path("~", "Downloads", "met_all.gz"))
```

## 2. **Check the dimensions, headers, footers. How many columns, rows are there?**

```{r}
dim(met)
head(met)
tail(met)
```

## 3. **Take a look at the variables.**

```{r}
str(met)
```

## 4. **Take a closer look at the key variables.**

```{r}
table(met$year)
table(met$day)
table(met$hour)
table(met$elev)
table(met$wind.sp)
met[met$elev==9999.0] <- NA
summary(met$elev)
met <- met[temp>-40]
met2 <- met[order(temp)]
head(met2)
```

## 5. **Check the data against an external data source.**

```{r}
met <- met[temp>-15]
met2 <- met[order(temp)]
head(met2)
```

## 6. **Calculate summary statistics**

```{r}
elev <- met[elev==max(elev), ]
summary(elev)
cor(elev$temp, elev$wind.sp, use="complete")
cor(elev$temp, elev$hour, use="complete")
cor(elev$wind.sp, elev$day, use="complete")
cor(elev$wind.sp, elev$hour, use="complete")
cor(elev$temp, elev$day, use="complete")
```

## 7. **Exploratory graphs**

```{r}
hist(met$elev, breaks=100)
hist(met$temp)
hist(met$wind.sp)

if(!require("leaflet")){ install.packages("leaflet")
library("leaflet")}
leaflet(elev) %>%
  addProviderTiles('OpenStreetMap') %>% 
  addCircles(lat=~lat,lng=~lon, opacity=1, fillOpacity=1, radius=100)

if(!require("lubridate")){ install.packages("lubridate")
library(lubridate)}
elev$date <- with(elev, ymd_h(paste(year, month, day, hour, sep= ' ')))
summary(elev$date)
elev <- elev[order(date)]
head(elev)

plot(elev$date, elev$temp, type='l')
plot(elev$date, elev$wind.sp, type='l')
```

## 8. Ask Questions

```{r}
cor(met2$elev,met2$temp,use='complete')

```

Is there a correlation between temperature and elevation?

There appears to be a somewhat moderate negative correlation between temperature and correlation in this case.